Automated neuroimaging data-management with Sea

### 2022-05-18 paper/sea-neuro/paper-neuro.tex

    The recent explosion in publicly available neuroimaging data has lead to new data
    management challenges, from storage infrastructure application processing times. To meet the storage, accessibility and
    security demands of neuroimaging data, large datasets have been stored in object stores provided by cloud storage providers.
    Standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids}, have been implemented to facilitate the sharing
    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide versioning and provenance capture of data.
    Furthermore, recent developments in neuroimaging pipelines have addressed computation time limitations by adopting machine-learning approaches.
    While all these solutions to Big Data-related data management exist, certain aspects, such as processing-related data-transfer overheads
    have received limited attention.

    The current largest neuroimaging datasets,
    the Human Connectome Project (HCP)~\cite{HCP} and the UK
    Biobank~\cite{ukbiobank}, reach up to Petabytes of data. For datasets of this
    volume, parallel high-capacity remote storage solutions are commonly relied
    upon as the persistent storage layer during processing. 
    Despite their respective high-scalability, these remote file systems cannot typically
    compete with the performance of compute-local storage (e.g. DRAM and SSDs). Should the pipelines 
    need to transfer large amounts of data to these remote storage locations, significant data
    processing overheads may occur.

    Big Data in neuroimaging can be present in two formats: 1) very large files
    (e.g. BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
    population of individuals, such as those found in MRI datasets, that
    collectively amount to very large datasets. Differences in the number and
    size of files result in different data transfer overheads. Whereas large
    file I/O is predominantly impacted by the underlying storage bandwidth,
    small file I/O results in a non-negligible latency overhead.

    Big Data frameworks have addressed the performance limitations arising from large data transfers via the
    implementation of two strategies: data locality and in-memory computing~\cite{zaharia2016apache, rocklin2015dask}.
    The collective aim of these strategies is to reduce overheads related to large data
    transfers by leveraging compute-local storage as cache.
    While these frameworks have been used
    to process neuroimaging data~\cite{rokem2021pan,thunder,boubela2016big},
    their popularity within the domain remains low due to efforts required to
    rewrite standardized neuroimaging toolboxes to utilize the frameworks.
    % Although it is possible to use the tools within the Big Data frameworks,
    % optimizations like in-memory computing would not be leveraged due to the
    % fact that neuroimaging tools are meant to be used as command-line tools
    % and do not provide interfaces that enable the data to be transferred
    % in-memory.

    Newer neuroimaging applications leveraging popular neuroimaging pipeline engines
    also do not benefit from processing-related data management.
    Although engines such as Nipype~\cite{nipype} and
    Joblib~\cite{joblib} do not prohibit the use of data-management
    strategies, they do not facilitate the integration of these strategies into their resulting workflow. To give
    neuroimaging applications data management capabilities, the applications
    must interact with a file system or library that enable the strategies.

    % Data management-specific scientific frameworks, such as
    % DataLad~\cite{halchenko2021datalad} aim to increase performance by focusing
    % on reproducibility, thereby reducing the need to recompute
    % results~\cite{wagner2021fairly}. We aim to complement frameworks such as
    % these, by reducing data transfer times when computation is required.

@@

    We present Sea~\cite{sea-comp}, a data-management library designed to reduce processing-related overheads. We view Sea as a complement
    to the existing stack of neuroimaging tools and standards meant to facilitate the processing of neuroimaging Big Data.    
    % \tristan{Intro logic is good but writing could be polished}
    % \tristan{Outlining the content of the paper would be useful}

>>>>

    Neuroimaging datasets continue to grow in size resulting in new data
    management challenges. The current largest neuroimaging datasets, such as
    the Human Connectome Project (HCP)~\cite{HCP} and the UK
    Biobank~\cite{ukbiobank} reach up to Petabytes of data. For datasets of this
    volume, there is no choice but to store the data in slower remote storage,
    which can incur important data transfer costs when processing. Should the
    processed data be of a similar magnitude in volume, application writes will
    also need to be made to compute-remote storage thus further prolonging
    application makespan.

    To minimize the costs of data transfers, Big Data frameworks, such as Apache
    Spark~\cite{zaharia2016apache} and Dask~\cite{rocklin2015dask} leverage
    memory and local storage as a cache. While these frameworks have been used
    to process neuroimaging data~\cite{rokem2021pan,thunder,boubela2016big},
    their popularity within the domain remains low due to the effort required to
    rewrite standardized neuroimaging toolboxes for the
    frameworks. 
    % Although it is possible to use the tools within the Big Data frameworks,
    % optimizations like in-memory computing would not be leveraged due to the
    % fact that neuroimaging tools are meant to be used as command-line tools
    % and do not provide interfaces that enable the data to be transferred
    % in-memory.

    Workflow engines used in neuroimaging, such as Nipype~\cite{nipype} and
    Joblib~\cite{joblib} instead focus on reducing compute time of workloads via
    parallelization of tasks.
    %This is because even with large datasets, neuroimaging data processing
    %remains split between compute and data intensive components.
    Although these workflow engines do not prohibit the use of data-management
    strategies, it is not directly integrated into the workflow. To give
    neuroimaging applications data management capabilities, the applications
    must interact with a file system that can do so.

    Data management-specific scientific frameworks, such as
    DataLad~\cite{halchenko2021datalad} aim to increase performance by focusing
    on reproducibility, thereby reducing the need to recompute
    results~\cite{wagner2021fairly}. We aim to complement frameworks such as
    these, by reducing data transfer times when computation is required.

    Big Data in neuroimaging can be present in two formats: 1) very large files
    (e.g. BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
    population of individuals, such as those found in MRI datasets, that
    collectively amount to very large datasets. Differences in the number and
    size of files result in different data transfer overheads. Whereas large
    file I/O is predominantly impacted by the underlying storage bandwidth,
    small file I/O results in a non-negligible latency overhead. Addressing both
    of these issues in neuroimaging is important, however, for this paper we
    focus on fMRI datasets, which are currently more likely to experience a
    latency overhead. We chose to focus on these types of datasets due to the
    number of well-established tools to process them.

@@

    In this paper, we evaluate Sea (Chapter~\ref{chapter:sea-comp}), a data-management library for HPC, on three neuroimaging preprocessing pipelines
    executed on three different datasets. We structure our analysis of Sea for the data-management of
    neuroimaging pipelines by first investigating whether Sea can effectively intercept the calls of these neuroimaging pipelines.
    We then investigate the presence of speedup by using Sea alongside neuroimaging pipelines and try to understand the conditions
    which result in greater speedups.

    % \tristan{Intro logic is good but writing could be polished}
    % \tristan{Outlining the content of the paper would be useful}
